---
title: "Vaccination and Public Health: Understanding the Correlation Between COVID-19 Vaccination Rates and Health Outcomes"
subtitle: "Assignment 4 ETC5512"
author: "Trisha Aurielle Andrea"
format: html
editor: visual
---

::: panel-tabset
<!-- Task 1 Tab: Data and Documenting -->

## Data and Documenting
### Topic of this work
How do COVID-19 vaccination rates correlate with COVID-19 case rates and mortality across different regions?

### Data download
```{r, echo=FALSE}
library(readr)
```

##### Downloading Dataset 1
###### Dataset link
https://www.kaggle.com/datasets/antgoldbloom/covid19-data-from-john-hopkins-university?resource=download

###### Description 
Global COVID-19 data, including the number of confirmed cases, fatalities, and recoveries for different nations and regions, are included in this collection. It is a trustworthy resource for monitoring the global spread of the pandemic and is updated frequently.

###### Download Process
The dataset can be downloaded directly from Kaggle by using the provided link.

###### Data Usage
The analysis of COVID-19's global impact and spread, as well as the correlation of vaccination rates with case and death data, depend on this data.

```{r, warning=FALSE}
covid_data_cases <- read_csv ("data/COVID-19 data from John Hopkins University/CONVENIENT_global_confirmed_cases.csv")
covid_data_deaths <- read_csv ("data/COVID-19 data from John Hopkins University/CONVENIENT_global_deaths.csv")
```

##### Downloading Dataset 2
###### Dataset link
https://www.kaggle.com/datasets/gpreda/covid-world-vaccination-progress
###### Description 
The number of people vaccinated against COVID-19 in various nations and regions is displayed in this dataset, which includes global vaccination statistics.  It also provides details on the coverage rate and vaccination manufacturers.

###### Download Process
The dataset can be downloaded directly from Kaggle by using the provided link.

###### Data Usage
Helpful in establishing a correlation between immunization progress and COVID-19 outcomes, such as recovery rates, fatalities, and cases.
```{r, warning=FALSE}
vaccination_data <- read_csv ("data/COVID-19 World Vaccination Progress/country_vaccinations.csv")
```

##### Downloading Dataset 3
###### Dataset link
https://www.kaggle.com/datasets/paultimothymooney/oxford-covid19-government-response-tracker

###### Description 
This dataset documents national governments' reactions to COVID-19, including travel bans, mask requirements, and lockdowns.  It contains daily information on the occurrence and severity of governmental actions.

###### Download Process
The dataset can be downloaded directly from Kaggle by using the provided link, and then i converted this xlsx file into csv format.

###### Data Usage
Analyzing the effects of government policy on vaccination rates, case counts, and mortality requires this data.  For a more thorough analysis, it can be paired with COVID-19 case data and vaccination data.

```{r, warning=FALSE}
Gov_response <- read_csv ("data/COVID-19 Government Response Tracker.csv")
```


<!-- Task 2 Tab: Writing your blogpost -->

## Blog post

### What's in this section

Here is where you should write your blogpost! (Task 2)

### Blogpost Structure

There is no strict structure for you to follow. However, here is a skeleton structure you may find helpful.

1.  Title (is set at the top of this document)
2.  Motivation\
3.  Data\
4.  Analysis\
5.  Conclusions\
6.  References

<!-- Task 3 Tab: Behind the Scenes -  -->

## Behind the Scenes

### What's in this section

Here is where you should tell us about your reflection on your analysis (Task 3).

Again, these are the details about **your** perspective and the gritty details behind the scenes of your analysis.
:::
