---
title: "Exploring the Correlation Between COVID-19 Vaccination Rates and Public Health Outcomes"
subtitle: "Assignment 4 ETC5512"
author: "Trisha Aurielle Andrea"
format: html
editor: visual
---

::: panel-tabset
<!-- Task 1 Tab: Data and Documenting -->

```{r, echo=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(DT)
library(Metrics)
```

## Data and Documenting

#### Topic 
#### Description
#### Dataset link
https://www.kaggle.com/datasets/omaebelfeki/graduate-admissions-dataset?select=Admission_Predict.csv

#### Download Process
#### Data Usage

```{r, warning=FALSE}
data <- read.csv("data/Admission_Predict.csv")
head(data)
```


<!-- Task 2 Tab: Writing your blogpost -->

## Blog post

#### Data Preview
```{r}
datatable(data)
```

#### Research Experience
##### Pie Chart of Research Experience
```{r}
research_data <- data.frame(
  Research = c(sum(data$Research == 1), sum(data$Research == 0)),
  Category = c("Research", "No Research")
)

ggplot(research_data, aes(x = "", y = Research, fill = Category)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  theme_void() +
  ggtitle("Distribution of Research Experience") +
  theme(legend.title = element_blank())
```
##### Research Experience vs Chance of Admit
```{r}
ggplot(data, aes(x = factor(Research), y = Chance.of.Admit)) +
  geom_bar(stat = "summary", fun = "mean", fill = "steelblue") +
  ggtitle("Research Experience vs Chance of Admit") +
  xlab("Research Experience (1 = Yes, 0 = No)") +
  ylab("Average Chance of Admit")

```


#### Bar Plot of University Rating vs Chance of Admit
```{r}
ggplot(data, aes(x = factor(University.Rating), y = Chance.of.Admit)) +
  geom_bar(stat = "summary", fun = "mean", fill = "orange") +
  ggtitle("University Rating vs Chance of Admit") +
  xlab("University Rating") +
  ylab("Average Chance of Admit")
```
#### Predicted vs. Actual Chance of Admit

```{r}
model <- lm(Chance.of.Admit ~ GRE.Score + TOEFL.Score + University.Rating + SOP + LOR + CGPA + Research, data = data)

predicted_values <- predict(model, data)

ggplot(data, aes(x = Chance.of.Admit, y = predicted_values)) +
  geom_point(color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  ggtitle("Predicted vs Actual Chance of Admit") +
  xlab("Actual Chance of Admit") +
  ylab("Predicted Chance of Admit") +
  theme_minimal()
```

#### Blogpost Structure

There is no strict structure for you to follow. However, here is a skeleton structure you may find helpful.

1.  Title (is set at the top of this document)
2.  Motivation\
3.  Data\
4.  Analysis\
5.  Conclusions\
6.  References

<!-- Task 3 Tab: Behind the Scenes -  -->

## Behind the Scenes

### What's in this section

Here is where you should tell us about your reflection on your analysis (Task 3).

Again, these are the details about **your** perspective and the gritty details behind the scenes of your analysis.
:::
